---
title: "Fun with R"
author: "Kristen Bystrom"
date: "September 21, 2019"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)
```

| Topics Covered                                        | Min | 
|-------------------------------------------------------|-----|
| Data Cleaning                                         | 10  | 
| Model Training                                        | 10  | 
| Intermediate Viualization                             | 10  | 
| Bonus Highlight: Learn how to make memes with R!      | 10  |

### Data Cleaning

Let's load in my fitbit data for the last month for activity and sleep.
```{r}
activity <- read.csv("Activity.csv", stringsAsFactors = FALSE)
sleep <- read.csv("Sleep.csv", stringsAsFactors = FALSE)
```

Let's take a look at the data.
```{r}
str(activity)
str(sleep)
```

Notice how several of the variables are stored as **Characters**. This is a problem because R thinks they are text values instead of dates or numbers.

Let's clean up the data types.

```{r, warnings = FALSE}
#install.packages("dplyr")
#install.packages("lubridate")
library(dplyr)
library(lubridate)

activity <- activity %>% mutate(
    Date = ymd(Date), 
    Calories.Burned = as.numeric(gsub(",", "", Calories.Burned)), 
    Steps = as.numeric(gsub(",", "", Steps)),
    Minutes.Sedentary = as.numeric(gsub(",", "", Minutes.Sedentary)),  
    Minutes.Lightly.Active = as.numeric(gsub(",", "", Minutes.Lightly.Active)),
    Minutes.Fairly.Active = as.numeric(gsub(",", "", Minutes.Fairly.Active)),
    Minutes.Very.Active = as.numeric(gsub(",", "", Minutes.Very.Active)),
    Activity.Calories = as.numeric(gsub(",", "", Activity.Calories))
)

sleep <- sleep %>% mutate(
    Start.Time = ymd_hm(Start.Time), 
    End.Time = ymd_hm(End.Time), 
    Minutes.REM.Sleep = as.numeric(gsub("/","", Minutes.REM.Sleep)),
    Minutes.Light.Sleep = as.numeric(gsub("/","", Minutes.Light.Sleep)),
    Minutes.Deep.Sleep = as.numeric(gsub("/","", Minutes.Deep.Sleep)),
    Date = date(End.Time)
    
)
```

If you look at the sleep data, there are some rows with 0 sleep recorded or less than two hours recorded. Let's assume these are nights where I charged my fitbit overnight and didn't record my sleep (and the less than 2 hours could be a nap). We want to get rid of these records since they will interfere with any models.

```{r}
sleep = sleep %>% filter(Minutes.Asleep > 120)
```

Similarily, in the activity data set, if there are 0 steps, the data probably didn't sync yet or I didn't wear my fitbit that day.

```{r}
activity = activity %>% filter(Steps > 0)
```

I want to be able to use my sleeping data with my activity data so I need to merge them together. 

```{r}
fitbit = inner_join(sleep, activity, by = "Date")
```


### Model Training

Now I have a clean (although small) data set. Let's build a simple model that predicts Calories burned based on some of the other information. There's not much you can do with a data set this small, but let's try something anyways to learn!

```{r}
library(caret)

inTraining <- createDataPartition(fitbit$Calories.Burned, p = .80, list = FALSE)
training <- fitbit[ inTraining,]
testing  <- fitbit[-inTraining,]
```


```{r}

set.seed(825)
bayesFit <- train(Calories.Burned ~ Minutes.Asleep + Minutes.Deep.Sleep + Steps + Minutes.Very.Active, data = training, 
                 method = "bayesglm")
bayesFit
```

Let's check how our model performs on our test set
```{r}
paste('Actual: ', testing$Calories.Burned)
paste('Predicted: ', predict(bayesFit, newdata = testing))
```


### Intermediate Visualization

Look how annoying 3D visualizations are to manipulate :p
```{r}
library(plotly)

plot_ly(data = fitbit, 
        x = ~Date, 
        y = ~Minutes.Very.Active,
        z = ~Calories.Burned)%>%
  add_markers() %>%
  layout(scene = list(xaxis = list(title = 'Date'),
                     yaxis = list(title = 'Minutes Active'),
                     zaxis = list(title = 'Calories Burned')))
```

Let's do better by using size, colour, etc to keep track of dimensions

```{r, warning = FALSE}
library(plotly)
library(broom)

m <- loess(Calories.Burned ~ Minutes.Very.Active, data = fitbit)

fitbit %>% 
  plot_ly(x = ~Minutes.Very.Active)%>%
  add_markers(y = ~Calories.Burned, size = ~Steps, text = fitbit$Date, showlegend = FALSE, name = "Day") %>%
  add_lines(y = ~fitted(loess(Calories.Burned ~ Minutes.Very.Active)),
            line = list(color = '#07A4B5'),
            name = "Loess Smoother", showlegend = TRUE)  %>%
  add_ribbons(data = augment(m),
              ymin = ~.fitted - 1.96 * .se.fit,
              ymax = ~.fitted + 1.96 * .se.fit,
              line = list(color = 'rgba(7, 164, 181, 0.05)'),
              fillcolor = 'rgba(7, 164, 181, 0.2)',
              name = "Standard Error")   %>%
  layout(xaxis = list(title = 'Minutes Very Active'),
         yaxis = list(title = 'Calories Burned'),
         legend = list(x = 0.80, y = 0.20))
```

### Bonus Highlight: Learn how to make memes with R!

The great thing about about source languages: anybody can develop a package (including you!). Here's one of my favourites!
https://cran.r-project.org/web/packages/meme/vignettes/meme.html


```{r}
#install.packages("meme")
library(meme)
```

```{r}
#Only need to run the following line if you are using Windows
if (.Platform$OS.type == "windows") {
    windowsFonts(
        Impact = windowsFont("Impact"),
        Courier = windowsFont("Courier")
    )
}

u <- system.file("success.jpg", package="meme")
myMeme <- meme(u, "went to an R Workshop","learned how to make memes with code")
myMeme
meme_save(myMeme, file="successR.png")
```


