---
title: "R Basics"
author: "Kristen Bystrom"
date: "September 21, 2019"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, collapse = TRUE)

#You may need to run install.packages("prettydoc") for this document to knit)
```

| Topics Covered                                        | Min | 
|-------------------------------------------------------|-----|
| Features of RStudio                                   | 5   | 
| Working with Packages and Libraries                   | 5   | 
| Understanding Functions                               | 5   | 
| Reading Data                                          | 5   |
| Beginner Visualization                                | 10  |
| Bonus Highlight: Learn how to make R rap to you!      | 10  | 


### Features of RStudio
- R/ R Studio/ Jupyter / R Studio Cloud
- Workspace
- Console
- Editor
- R Markdown
- Running Code Chunks
- Environment
- Files/ Plots/ Packages/ **Help**/ Viewer


### Working with Packages and Libraries  

When you install R, it comes with some basic capabilities. 
However, R is an **open source** language which means that anybody can make **packages** that add extra features you can use. There are two steps to use these packages.

1. Install the **package** onto your computer (you only need to do this once)
```{r}
#install.packages("ggplot2")
```

2. Load the **library** in your current R session (you need to do this for every new R program you write and should have this code in your library)
```{r}
library(ggplot2)
```


### Understanding Functions    

You have already seen **functions**! They look like a word followed by a set of soft brackets like library(), read.csv(), sum(), or c(). The word before the soft brackets describes what the function does, and you may need to add **arguments** inside the brackets to help give the expressions information they need to operate. 

Let's try some functions!

The c() function creates a vector which can store one kind of data so you can use it later. The arguments it takes are data observations
```{r}
myNumericVector <- c(2,3,7,8,10)
myCharacterVector <- c("Blue", "Blue", "Red", "Red", "Red")

myNumericVector
myCharacterVector
```

The sum() function adds up all the arguments
```{r}
sum(2, 2)
sum(2,3,7,8,10)
sum(myNumericVector)
```


### Reading Data      
One of the main ways people use R is to work with data to do visualization or statistical analysis. To do this, it's best to put the data file in the same folder as your **R Project**, and then use a function to read it in to your R Session.

```{r}
myData <- read.csv("totally_normal_data.csv", header = FALSE)
```

The head() function is helpful to see what the data looks like. So is the str() function.

```{r}
head(myData, n = 10)
str(myData)
```

Rename the columns to something descriptive.

```{r}
colnames(myData) <- c("Beyonce", "Khaled")
```

### Beginner Visualization     

Let's see what the data looks like. Remember, you may need to **install** the **package** first before **loading** the **library**. 
```{r}
# install.packages("ggplot2")

library(ggplot2)

ggplot(myData, aes(x = Beyonce, y = Khaled))+
  geom_point()

```

Credit for the dataset to:
https://cran.r-project.org/web/packages/datasauRus/vignettes/Datasaurus.html

We can do a lot better than that!

```{r}
# install.packages("ggplot2")

library(ggplot2)
library(ggthemes)
library(viridis)

ggplot(myData, aes(x = Beyonce, y = Khaled, color = Khaled))+
  geom_point()+
  ggtitle("Land Before Time")+
  xlab("Dinosaur Height")+
  ylab("Dinosaur Width")+
  theme_few()+scale_color_viridis()

```

### Bonus Highlight: Learn how to make R rap to you! 

So why is R the best programming language? Well let me show you a few of my favourite things:

I often use the beepr package in R which has handy ringtones and short songs that can play when your code finishes executing.

However, R also has a library called BRRR
```{r}
if(!require(devtools)) {install.packages('devtools')}
devtools::install_github("brooke-watson/BRRR")
library(BRRR)
```

Let's test it!

```{r}
skrrrahh(sound = 26)
```

```{r}
skrrrahh("snoop")
```

```{r}
skrrrahh(10)
```

You can also change your options so DJ Kaled consoles you every time you hit an error message
```{r}
options(error = function() {skrrrahh(34)})
```